Trafficking synaptic cargo in complex morphologies involves tradeoffs in speed and precision

### 2015-08-22 latex/paper.tex
fix typos, add some figure legends

\author[1,2,3,*]{Alex H. Williams}
\author[2]{Cian O'Donnell}
\author[4]{Eve Marder}
\author[2,5]{Terrence Sejnowski}
\author[4]{Timothy O'Leary}

\affil[1]{Department of Neurosciences, University of California, San Diego, La Jolla, CA 92093, USA}
\affil[2]{Howard Hughes Medical Institute, Salk Institute for Biological Studies, La Jolla, CA 92037, USA}
\affil[3]{Department of Neurobiology, Stanford University, Stanford, CA 94305, USA}
\affil[4]{Volen Center and Biology Department, Brandeis University, Waltham, MA 02454, USA}
\affil[5]{Division of Biological Sciences, University of California at San Diego, La Jolla, CA 92093, USA}
\affil[*]{Address correspondence to: ahwillia@stanford.edu}

@@

@@

\begin{figure}[!b]
\begin{center}
\includegraphics[width=0.9\columnwidth]{00_stochastic.png}
\caption{A simple mass-action model of dendritic transport}
\end{center}
\end{figure}

@@

To obtain a general model that can accommodate variations in the biophysical details, we consider microtubule-based transport as a biased random walk along a one-dimensional cable \citep{Bressloff_2006,Newby_2010,Bressloff_2009}.
For each time step, the cargo moves left, right, or remains the in the same place.
In the simplest model, the probabilities associated with each movement are fixed and independent across each time step (Fig. 1B, top panel).
A potentially more realistic model incorporates \textit{hysteresis} into the biased walk (see Methods). In this case, the cargo is more likely to continue in the same direction as it moved on the previous time step (Fig. 1B, bottom panel; see, e.g., \cite{Soundararajan_2014}).

@@

\hl{MOVE THIS PARAGRAPH TO METHODS/DISCUSSION?} A central assumption of this model is that the movement of molecules between neighboring compartments is a stochastic and memoryless process. This assumption is reasonable if the run length of molecular motors is comparable to the size of the compartments; this appears to be the case under many \citep{Muller_2008,Verbrugge_2009}, thought not necessarily all \citep{Dynes_2007} circumstances (see Discussion). In simulations, we observed that this model was not very sensitive to violations of this assumption; the population dynamics of cargoes with substantial run lengths could still be well-fit by this class of models (Fig. 1F, lower panel; Supplementary Fig?).

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\columnwidth]{01_mass_action.png}
\caption{Delivery of cargo to active synapses in a complex dendritic tree}
\end{center}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\columnwidth]{03_convergence.png}
\caption{Bottlenecks in transport}
\end{center}
\end{figure}

@@

One experimentally testable prediction of this analysis is that transport to distal compartments should be quickened by allowing higher levels of $u$ into the proximal compartments at steady-state (Fig. 3E).
This might be tested experimentally by characterizing the convergence time of a retrogradely transported molecule that aggregates at recently activated synapses, such as \textit{Arc} mRNA (\citet{Steward_1998}, see discussion).
To illustrate this in the model CA1 cell, we characterized the time course of transport to the distal apical dendrite (stimulating stratum lacunosum/moleculare), proximal apical dendrite (stimulating stratum radiatum), entire apical dendrite (stimulating both layers), and entire cell (seizure). Interestingly, the model converged more slowly to distal input alone, than to paired distal and proximal input, or to an entirely uniform distribution (Fig. 3F).

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\columnwidth]{04_sushi_belt.png}
\caption{Multiple strategies for transport in a model including nonuniform microtubular detachment/activation}
\end{center}
\end{figure}

One way to analyze this system is to assume that these three processes --- trafficking, detachment, and degradation --- occur on separate timescales.
If trafficking is sufficiently faster than detachment ($a,b \gg c$), then $u$ approaches a quasi-steady state distribution defined by our previous analysis (equation 5).
We then choose detachment rate constants that transform the microtubular distribution into our desired distribution for $u^\star$:
\begin{equation}
c_i \propto \frac{\tilde{u}^\star_i}{\tilde{u}_i}
\end{equation}
Here, $\tilde{u}$ and $\tilde{u}^\star$ respectively denote the quasi-steady state distributions for $u$ and $u^\star$, respectively.
Due to the degradation reaction, the system no longer approaches a steady-state; however, as long as degradation is sufficiently slow ($c \gg d$) the desired distribution is transiently achieved.

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\columnwidth]{05_heterosynaptic_ca1.png}
\caption{Heterosynaptic and non-local interactions.
}
\end{center}
\end{figure}

@@

In figure 5, we showed that scaling the detachment rates ($c_i$), and leaving the transport rates ($a_i$,$b_i$) fixed produces a proximal bias in cargo delivery.
That is, synapses closer to the soma are more likely to capture cargo at the expense of distal synapses.
We reasoned that increasing the anterograde transport rate of cargo could correct for this bias, producing transport rules that are both fast and precise.

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\columnwidth]{06_heterosynaptic_cable.png}
\caption{Tradeoff between speed of consolidation and specificity of potentiation.}
\end{center}
\end{figure}

>>>>

\author[1,2,3,*]{Alex H. Williams}
\author[2]{Cian O'Donnell}
\author[4]{Eve Marder}
\author[2,5]{Terrence Sejnowski}
\author[4,*]{Timothy O'Leary}

\affil[1]{Department of Neurosciences, University of California, San Diego, La Jolla, CA 92093, USA}
\affil[2]{Howard Hughes Medical Institute, Salk Institute for Biological Studies, La Jolla, CA 92037, USA}
\affil[3]{Department of Neurobiology, Stanford University, Stanford, CA 94305, USA}
\affil[4]{Volen Center and Biology Department, Brandeis University, Waltham, MA 02454, USA}
\affil[5]{Division of Biological Sciences, University of California at San Diego, La Jolla, CA 92093, USA}
\affil[*]{Address correspondence to: ahwillia@stanford.edu, toleary@brandeis.edu}

@@

\hl{}

@@

\begin{figure}[!tb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{00_stochastic.png}
\caption{A simple mass-action model of dendritic transport}
\end{center}
\end{figure}

@@

To obtain a general model that can accommodate variations in the biophysical details, we consider microtubule-based transport as a biased random walk along a one-dimensional cable \citep{Bressloff_2006,Newby_2010,Bressloff_2009}.
For each time step, the cargo moves left, right, or remains in the same place.
In the simplest model, the probabilities associated with each movement are fixed and independent across each time step (Fig. 1B, top panel).
A potentially more realistic model incorporates \textit{hysteresis} into the biased walk (see Methods). In this case, the cargo is more likely to continue in the same direction as it moved on the previous time step (Fig. 1B, bottom panel; see, e.g., \cite{Soundararajan_2014}).

@@

\hl{MOVE THIS PARAGRAPH TO METHODS/DISCUSSION?} A central assumption of this model is that the movement of molecules between neighboring compartments is a stochastic and memoryless process. This assumption is reasonable if the \hl{run length (jargon?)} of molecular motors is comparable to the size of the compartments; this appears to be the case under many \citep{Muller_2008,Verbrugge_2009}, thought not necessarily all \citep{Dynes_2007} circumstances (see Discussion). In simulations, we observed that this model was not very sensitive to violations of this assumption; the population dynamics of cargoes with substantial run lengths could still be well-fit by this class of models (Fig. 1F, lower panel; Supplementary Fig?).

@@

\begin{figure}[!tb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{01_mass_action.png}
\caption{Local trafficking parameters determine the spatial distribution of biomolecules by a simple rule.
\textbf{(A)} Diagram of the mass action transport model for a simple branched morphology.
%The steady-state ratio of $u$ between any two neighboring compartments is equal to the reciprocal ratio of the trafficking rate constants between those compartments.
\textbf{(B)} A simulation of a nine compartment cable, with trafficking rate constants set to produce a linear gradient using the steady-state relation shown in panel A.
%The initial distribution of $u$ is uniform (t=0); over time the system achieves the desired linear gradient (steady-state occurs around t=5, in this example).
\textbf{(C)} The slope of the linear gradient shown in panel B is determined by the total amount of $u$ present in the model; the slope increases as more of $u$ is present, but the linear profile is preserved.
\textbf{(D)} A model of a CA1 pyramidal cell with 742 compartments based on \cite{Migliore_2012}; excitatory synapses were added at the locations marked by red dots. 
\textbf{(E)} The average membrane potential in each compartment of the CA1 model cell was used to define a desired spatial distribution, and transport rate constants were chosen according to equation (4). The desired spatial distribution of cargo emerges over time, after starting in the soma.
\textbf{(F)} The desired and steady-state profile of $u$ for each compartment in the simulation shown in panel E.
}
\end{center}
\end{figure}

@@

\begin{figure}[!tb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{03_convergence.png}
\caption{ Convergence to steady-state is slow when molecules must be transported across bottlenecks ---compartments with low target levels.
\textbf{(A)} A three-compartment transport model, with the middle compartment acting as a bottleneck. The vertical bars represent the desired steady-state level of $u$ in each compartment. The rate of transport into the middle compartment is small ($\epsilon$, dashed harpoons) relative to transport out of the middle compartment.
\textbf{(B)} As $\epsilon$ decreases, the model converges more slowly and the steady-state level decreases in the middle compartment.
\textbf{(C)} Simulations (black dots) confirm that the time to convergence is given by the smallest non-zero eigenvalue of the system (analytically calculated line). This eigenvalue can be thought of as the rate-limiting step of the system.
\textbf{(D)} Convergence ($L_1$ distance) to a uniform target distribution (red line) is faster than a target distribution containing a bottleneck (blue line) in the CA1 model.
\textbf{(E)} For all compartments that reach a threshold level ($u_i = 0.001$), the simulated time it takes to reach threshold is plotted against the distance of that compartment to the soma.
\textbf{(F)} Predicted convergence times for various target distributions (str. rad., stratum radiatum; str. lac., stratum lacunosum/moleculare).}
\end{center}
\end{figure}

@@

One experimentally testable prediction of this analysis is that transport to distal compartments should be quickened by allowing higher levels of $u$ into the proximal compartments at steady-state (Fig. 3E).
This might be tested by characterizing the convergence time of a retrogradely transported molecule that aggregates at recently activated synapses, such as \textit{Arc} mRNA (\citet{Steward_1998}, see discussion).
To illustrate this in the model CA1 cell, we characterized the time course of transport to the distal apical dendrite (stimulating stratum lacunosum/moleculare), proximal apical dendrite (stimulating stratum radiatum), entire apical dendrite (stimulating both layers), and entire cell (seizure). Interestingly, the model converged more slowly to distal input alone, than to paired distal and proximal input, or to an entirely uniform distribution (Fig. 3F).

@@

\begin{figure}[!tb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{04_sushi_belt.png}
\caption{Multiple strategies for transport in a model including nonuniform microtubular detachment/activation}
\end{center}
\end{figure}

One way to analyze this system is to assume that these three processes --- trafficking, detachment, and degradation --- occur on separate timescales.
If trafficking is sufficiently faster than detachment ($a,b \gg c$), then $u$ approaches a quasi-steady state distribution defined by our previous analysis (equation 4).
We then choose detachment rate constants that transform the microtubular distribution into our desired distribution for $u^\star$:
\begin{equation}
c_i \propto \frac{\tilde{u}^\star_i}{\tilde{u}_i}
\end{equation}
Here, $\tilde{u}$ and $\tilde{u}^\star$ respectively denote the quasi-steady state distributions for $u$ and $u^\star$, respectively.
Due to the degradation reaction, the system no longer approaches a steady-state; however, as long as degradation is sufficiently slow ($c \gg d$) the desired distribution is transiently achieved.

@@

\begin{figure}[!tb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{05_heterosynaptic_ca1.png}
\caption{Heterosynaptic and non-local interactions.
}
\end{center}
\end{figure}

@@

\hl{have detachment rate be tunable by amount of stuff present???}

In figure 5, we showed that scaling the detachment rates ($c_i$), while leaving the transport rates ($a_i$, $b_i$) fixed produces a proximal bias in cargo delivery.
That is, synapses closer to the soma are more likely to capture cargo at the expense of distal synapses.
We reasoned that increasing the anterograde transport rate of cargo could correct for this bias, producing transport rules that are both fast and precise.

@@

\begin{figure}[!tb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{06_heterosynaptic_cable.png}
\caption{Tradeoff between speed of consolidation and specificity of potentiation.}
\end{center}
\end{figure}
